Eliminating prior-bias from sparse-projection tomographic reconstructions

### 2020-07-17 paper/tci.tex
Made a few changes suggested by Prof.Sharat

\bibliographystyle{IEEEtran}
\usepackage{lineno,hyperref}
\usepackage{amsmath,graphicx, amssymb,tabularx,booktabs}
\usepackage{subcaption}
\usepackage[font=small,skip=2pt]{caption}
\usepackage{boxedminipage}
%\usepackage{algpseudocode}
%\usepackage{algorithm}
\usepackage{xcolor}
%\usepackage[sort&compress]{natbib}

@@

  If the goal is to track the changes while simultaneously reducing
  sub-sampling artefacts, we propose (1) acquiring measurements from
  an \emph{extremely small} number of views and using a global
  \emph{`uniform'} prior-based reconstruction. If the goal is to
  observe \textbf{details} of new changes, we propose (2) acquiring
  measurements from a \emph{moderate} number of views (albeit, still
  sub-Nyquist), and using a more involved reconstruction routine. We
  show that in the latter case, a `spatially-varying' technique is
  appropriate in order to prevent the prior from adversely affecting the
  reconstruction of new structures that are absent in any of the
  earlier scans. The reconstruction of new regions is safeguarded from
  the bias of the prior by computing regional weights that moderate
  the local influence of the priors. We are thus able to effectively
  reconstruct both the old and the new structures in the test. We have
  tested the efficacy of our method on synthetic as well as real
  projection data. The results demonstrate the use of both uniform and
  spatially-varying priors in different scenarios. Our methods
  significantly improve the overall quality of the reconstructed data
  while minimizing the number of measurements needed for imaging in
  longitudinal studies.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
Computed Tomography (CT) deals with the recovery of an entire object
from a limited set of projection data which are acquired by passing
X-rays at different orientations (`views'). It is preferable to
minimize the radiation exposed in order to prevent any potential
damage to it and in order to reduce the acquisition time. Therefore,
current research seeks to either significantly reduce the radiation
intensity required to reconstruct with adequate
fidelity~\cite{yang2018,Lin2016,Xie2017,gopal2019low} or significantly
reduce the number of measurements required to reconstruct with
adequate fidelity. For the latter case, there are two lines of
pursuit. One is to intelligently choose those sets of projection views
that capture most
information~\cite{King2018,Anthony2018,barkan17,fischer16,andrei14},
and the other, which is the focus of this paper, is to design the
reconstruction algorithm in order to achieve the most accurate
recovery of the underlying slice, given the measurements from any
limited set of views~\cite{yang2018,geyer2015,kilic2011}.

In conventional data acquisition techniques, the tomographic
measurements $\boldsymbol{y}$ are acquired by sampling the physical
object $\boldsymbol{x}$ uniformly with a substantial number of views,
ideally above the Nyquist rate. In such a case when there are
sufficient measurements, reconstruction using the conventional
filtered backprojection (FBP) suffices, as seen in
Fig.~\ref{fig:story} (first column, 600 views). The figure shows the
ground truth image (260x260) of naturally growing sprouts at the top
left (the details of the dataset is postponed to
Section~\ref{sec:results_uniform_prior}).

@@

 \begin{figure}[t]
\centering
	\includegraphics[width=0.4\textwidth]{../images/story/story.png}
        \caption{Illustration of the use of various priors as the
          number of views reduces. For extremely large number of views
          (600 views), the FBP reconstruction (first column) is of
          very good quality (Structural Similarity Index Metric,
          SSIM=0.92). As the number of views become limited (100
          views), the reconstruction using sparsity prior (middle
          column) is better both visually and quantitatively than FBP
          (SSIM of 0.90 vs 0.83). If the number of views is
          drastically reduced (30 views), the presence of both
          sparsity prior and object prior improves the reconstruction
          (SSIM = 0.85) when compared to presence of only sparsity
          prior (SSIM = 0.82) or no prior (SSIM = 0.69). In all cases,
          the ground truth (260x260) appears in the top left.}
 \label{fig:story}
 \end{figure}

 When the number of views is significantly reduced, the sparsity prior
 alone is not sufficient.  In such cases additional information
 specific to the object being scanned (i.e., object prior) is useful
 in further improving the reconstruction.  While this has been done in
 the context of dynamic CT-scans (as discussed in
 Section~\ref{sec:related}), in this paper we use a set $S$ of
 previous scans of the same object that is more appropriate in a
 longitudinal context. As seen in Fig.~\ref{fig:story} (last column,
 30 views) our method uses the prior scans \emph{uniformly} by
 creating an eigenspace representation
 (Section~\ref{sec:method_uniform_prior}).

 However, with the use of such an object prior, a new challenge
 emerges. The prior set S may potentially overwhelm the necessary
 details, or since several prior scans are available, finding the
 right prior may be an issue. We therefore seek an algorithm that
 estimates the location and magnitude of new changes in the (unknown)
 test. As we show in Section~\ref{sec:method_spatially_varying_prior},
 this eventually prevents the prior from adversely affecting the
 reconstruction of new regions in the test. We refer to this method as
 a \textit{spatially-varying} global prior-based reconstruction
 routine, that still uses all the previous scans in the set $S$.

@@

We also demonstrate how the global prior can impose an inflexible
constant weight (and hence an unnecessary bias) when reconstructing
the data. As a solution, we present a method to moderate the control
of the prior by estimating and imposing spatially-varying weights to
the prior in order to reconstruct new structures accurately. This
spatially-varying prior tunes the effect of the templates in different
regions of the reconstruction.

@@

\section{Application: Reconstruction for CT-guided radio-ablation study}
\label{sec:tmh}
Before diving into the details of the uniform and the
spatially-varying global prior methods, we first show how both the
techniques can be applied to our advantage in a real-life medical
longitudinal study. Here, our data~\footnote{Source: Tata Memorial
  Centre~\cite{tmh}, Parel, Mumbai.  This is the national
  comprehensive centre for the prevention, treatment, education and
  research in cancer, and is recognized as one of the leading cancer
  centres in India.} consists of successive scans of the liver taken
during a radio-frequency ablation procedure. In such a
procedure~\cite{Dong2015}, the physician inserts a thin needle-like
probe into the organ. Repeated CT scans of the patient need to be
acquired in order to track the movement of the needle and to ensure
that it is reaching the appropriate target tumor. Once the needle hits
the tumor, a high-frequency electric current is passed through the tip
of the probe and this burns the malignant tumor (ablation). The scan
at this ablation stage must also reveal accurate details of the
changes within. In this context, we classify the goal of any of our
reconstruction techniques into two categories: 
\begin{enumerate}
\item To track the position of the needle in a relatively well-reconstructed background.%Initially, when our goal is to only track the location of the probe while simultaneously reducing sub-sampling artefacts, we advocate capturing measurements from a small number of views (`few-view' imaging) and using uniform prior based reconstruction.
\item To accurately observe the new changes amidst a relatively well reconstructed background after the needle touches the tumor.%Later, when the probe is proximal to or touches the tumor, our goal is to observe details of tumor ablation while simultaneously reducing sub-sampling artefacts. In this case, we propose to capture measurements from a slightly higher number of views (`moderate-view' imaging) to acquire more information, and use a spatially-varying prior-based reconstruction. The spatially-varying technique moderates the effect of the prior in the reconstruction of new changes in the object being scanned.
\end{enumerate}

%--------------------------Results on TMH data--------------------------
%------------------RFA 2 data
\begin{figure}[h!]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template1.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 1}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template2.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 2}
    \end{subfigure}
     \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template3.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 3}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template4.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 4}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template5.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 5}
    \end{subfigure}
              \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template6_color.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 6}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template7.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 7}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/colorTestIm_ablation.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 8}
    \end{subfigure}     
     \caption{Radio-frequency ablation dataset: One of the slices (512 x 512) from each of the 8 scan volumes of a longitudinal study dataset of the liver. Note that in volumes (a) through (g), the needle (shown in red in (f)) approaches the target tumor. (h) the organ after the ablation: this slice is displayed on a separate intensity scale to enable proper viewing of the region marked in green that shows the after-effects of ablating the tumor.} 
\label{fig:RFA2_test_object-prior}
\end{figure}

%--------------------RFA 2 few views
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{white}{green}{\includegraphics[width=0.23\columnwidth]{../images/tmh/RFA2/very_few_views/colorTestIm.png}}}\hfill
\subcaptionbox{Backprojection}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/fbp.png}}\hfill
\subcaptionbox{l1-ls}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/cs.png}}\hfill
\subcaptionbox{Proposed\\ method: uniform prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/plain_pca.png}}
\caption[Representative results-1]{\textbf{Goal: Track new changes.}\small{ Reconstruction of slice 7 (`test') of Fig.~\ref{fig:RFA2_test_object-prior} from only 90 views, using (b) Filtered backprojection and no prior resulting in streaks, SSIM $= 0.48$ (c) l1-ls resulting in blurred bone structures, SSIM $=0.35$ and (d) uniform global prior (slices 1-6 of Fig.~\ref{fig:RFA2_test_object-prior} are used as object-prior) resulting in clear bone structures with less streaks, SSIM $=0.55$. The region enclosed in red rectangle is our RoI as it contains both the new position of the needle and some background. All SSIM values are computed for this RoI.}}
\label{fig:RFA2_very_few_views}
\end{figure}

@@

\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{white}{blue}{\includegraphics[width=0.23\columnwidth]{../images/tmh/RFA2/few_views/colorTestIm.png}}}\hfill
\subcaptionbox{Backprojection}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/fbp.png}}\hfill
\subcaptionbox{l1-ls}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/cs_dct.png}}\hfill
\subcaptionbox{Proposed\\ method: uniform\\ prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/plain_pca.png}}
\subcaptionbox{Proposed\\ method: spatially-varying prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/weighted_pca_all_methods_kk_0_01.png}}
\subcaptionbox{Weights map}{\includegraphics[width=0.275\columnwidth]{../images/tmh/RFA2/few_views/weights_colormap.png}}
\caption[Representative results-2]{\textbf{Goal: Observe details of new changes.} Reconstruction of slice 8 of Fig.~\ref{fig:RFA2_test_object-prior} from 120 views, using (b) FBP with SSIM $= 0.50$ (c) l1-ls with SSIM $=0.46$ and (d) uniform global prior, with SSIM $=0.51$ (\emph{notice dominance of the prior: a prominent residual shadow of the needle which was present in the object-prior, but not present in the test image}), and (e) spatially-varying prior with SSIM $=0.56$ (\emph{notice that the dominance of the prior is significantly controlled}). The region enclosed in green rectangle is our Region of Interest (RoI) as it contains both the new position of the needle and some background. The SSIM is computed in this RoI. (f) shows the computed weights map (defined later in the paper) used for reconstruction. Darker intensities indicate lower weights to prior as these are the regions of new changes.}
\label{fig:RFA2_few_views}
\end{figure}

Here, we demonstrate the combined use of l1-ls (CS) with the global
prior, in two flavors: the vanilla (uniform) and spatially-varying
global prior-based reconstruction. The choice of the number of
measurement views and the type of reconstruction -- uniform or
spatially-varying, is driven by the goal of the procedure. First, in
order to track the needle, a very small number of views is sufficient
because the needle has a very high attenuation coefficient when
compared to that of the organs. We use the uniform global prior
reconstruction here to reduce the artefacts due to sub-sampling. The
uniform method is fast and sufficient to track the position of the
needle.  Once the needle reaches the site of the tumor, we propose
changing the imaging protocol to acquire measurements from a moderate
number of views. This will enable us to get more information about the
new changes. In addition, we then deploy the spatially-varying prior
method in order to locate the regions of new changes and penalize any
dominance of the prior in these regions.  Regardless of the imaging
protocol we use (`few' or `moderate'), the number of views is smaller
(atleast one-fifth) than the conventional number of views used in a
standard hospital setting.

The dataset from this longitudinal medical study consists of 8 scans
taken during the ablation procedure. We demonstrate our method for 2D
reconstruction by choosing a single slice from each of the 8 volumes
as our dataset. Note that all these 8 slices are located at the
\emph{same} index~\footnote{The notion of~\textit{same index} (slice
  number corresponding to the same depth) makes sense in the context,
  because in such problems, the different scans are aligned with each
  other.} within each of the respective
volumes. Fig.~\ref{fig:RFA2_test_object-prior} shows the chosen set of
2D slices (each of size $512 \times 512$) from the different
volumes. Observe that the needle is seen in all of the first 7 slices
and the effect of ablation is seen in the 8th slice. \\

@@

\textbf{Observing details of the ablation:} Next, we choose slices 1-7 as our object-prior and reconstruct slice 8 from 120 views i.e. a somewhat higher number of views this time. Fig.~\ref{fig:RFA2_few_views} shows the reconstructions of slice 8 by different methods. We see that the spatially-varying prior reconstruction brings in the advantage of the prior without it adversely affecting the new regions.

%---------------------------------Uniform section---------------------
\section{Method: uniform prior-based reconstruction}
\label{sec:method_uniform_prior}
Having presented the application, we first review the algorithm~\cite{my_dicta_paper} for a uniform (global) prior-based reconstruction. Principal Component Analysis (PCA) has been traditionally used to find the significant modes of (Gaussian corrupted) data. In this regard, it has been widely applied in the context of data compression. However, PCA can also be seen as a tool to provide an orthogonal basis to represent the space in which most of the test data could lie (except the new changes). This space is constructed from the available set of previously scanned objects which must cover a realistically representative range of structures. %, which was shown~\cite{my_dicta_paper} to be better when compared to dictionary-based priors. %A preliminary version of this particular method was earlier presented in \cite{my_dicta_paper}.

>>>>

\bibliographystyle{IEEEtran}
\usepackage{lineno,hyperref}
\usepackage{amsmath,graphicx, amssymb,tabularx,booktabs}
\usepackage{subcaption}
\usepackage[font=small,skip=2pt]{caption}
\usepackage{boxedminipage}
%\usepackage{algpseudocode}
%\usepackage{algorithm}
\usepackage{xcolor}
%\usepackage[sort&compress]{natbib}
\usepackage{footnote}
\makesavenoteenv{tabular}

@@

  If the goal is to track the changes while simultaneously reducing
  sub-sampling artefacts, we propose (1) acquiring measurements from
  an \emph{extremely small} number of views and using a
  \emph{`uniform'} prior-based reconstruction. If the goal is to
  observe \textbf{details} of new changes, we propose (2) acquiring
  measurements from a \emph{moderate} number of views (albeit, still
  sub-Nyquist), and using a more involved reconstruction routine. We
  show that in the latter case, a `spatially-varying' technique is
  appropriate in order to prevent the prior from adversely affecting the
  reconstruction of new structures that are absent in any of the
  earlier scans. The reconstruction of new regions is safeguarded from
  the bias of the prior by computing regional weights that moderate
  the local influence of the priors. We are thus able to effectively
  reconstruct both the old and the new structures in the test. We have
  tested the efficacy of our method on synthetic as well as real
  projection data. The results demonstrate the use of both uniform and
  spatially-varying priors in different scenarios. Our methods
  significantly improve the overall quality of the reconstructed data
  while minimizing the number of measurements needed for imaging in
  longitudinal studies.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
Computed Tomography (CT) deals with the recovery of details of an
object's interior
from a limited set of projection data which are acquired by passing
X-rays at different orientations (`views'). It is preferable to
minimize the radiation exposed in order to prevent any potential
damage to it and in order to reduce the acquisition time. Therefore,
current research seeks to either significantly reduce the radiation
intensity required to reconstruct with adequate
fidelity~\cite{yang2018,Lin2016,Xie2017,gopal2019low} or significantly
reduce the number of measurements required to reconstruct with
adequate fidelity. For the latter case, there are two lines of
pursuit. One is to intelligently choose those sets of projection views
that capture most
information~\cite{King2018,Anthony2018,barkan17,fischer16,andrei14},
and the other, which is the focus of this paper, is to design the
reconstruction algorithm in order to achieve the most accurate
recovery of the underlying slice, given the measurements from any
limited set of views~\cite{yang2018,geyer2015,kilic2011}.

In conventional data acquisition techniques, the tomographic
measurements $\boldsymbol{y}$ are acquired by sampling the physical
object $\boldsymbol{x}$ uniformly with a substantial number of views,
ideally above the Nyquist rate. In such a case when there are
sufficient measurements, reconstruction using the conventional
filtered backprojection (FBP) suffices, as seen in
Fig.~\ref{fig:story} (first column, 600 views). The figure shows the
ground truth image (260$\times$260) of naturally growing sprouts at the top
left (the details of the dataset is postponed to
Section~\ref{sec:results_uniform_prior}).

@@

 \begin{figure}[t]
\centering
	\includegraphics[width=0.45\textwidth]{../images/story/story.png}
        \caption{Illustration of the use of various priors as the
          number of views reduces. For extremely large number of views
          (600 views), the FBP reconstruction (first column) is of
          very good quality (Structural Similarity Index Metric,
          SSIM=0.92). As the number of views become limited (100
          views), the reconstruction using sparsity prior (middle
          column) is better both visually and quantitatively than FBP
          (SSIM of 0.90 vs 0.83). If the number of views is
          drastically reduced (30 views), the presence of both
          sparsity prior and object prior improves the reconstruction
          (SSIM = 0.85) when compared to presence of only sparsity
          prior (SSIM = 0.82) or no prior (SSIM = 0.69). In all cases,
          the ground truth (260x260) appears in the top left.}
 \label{fig:story}
 \end{figure}

 When the number of views is significantly reduced, the sparsity prior
 alone is not sufficient.  In such cases additional information
 specific to the object being scanned (i.e., object prior) is useful
 in further improving the reconstruction.  While this has been done in
 the context of dynamic CT-scans (as discussed in
 Section~\ref{sec:related}), in this paper we use a set $S$ of
 previous scans of the same object that is more appropriate in a
 longitudinal context. Specifically, longitudinal studies here refers
 to the acquisition of sequential CT scans of the same subject to
 track time-evolving changes within the subject's interior. As seen in
 Fig.~\ref{fig:story} (last column, 30 views) our method uses the
 prior scans \emph{uniformly} by creating an eigenspace representation
 (Section~\ref{sec:method_uniform_prior}).

 However, with the use of such an object prior, a new challenge
 emerges. The prior set S may potentially overwhelm the necessary
 details, or since several prior scans are available, finding the
 right prior may be an issue. We therefore seek an algorithm that
 estimates the location and magnitude of new changes in the (unknown)
 test. As we show in Section~\ref{sec:method_spatially_varying_prior},
 this eventually prevents the prior from adversely affecting the
 reconstruction of new regions in the test. We refer to this method as
 a \textit{spatially-varying} prior-based reconstruction
 routine, that still uses \textit{all} the previous scans in the set $S$.

@@

We also demonstrate how the uniform prior can impose an inflexible
constant weight (and hence an unnecessary bias) when reconstructing
the data. As a solution, we present a method to moderate the control
of the prior by estimating and imposing spatially-varying weights to
the prior in order to reconstruct new structures accurately. This
spatially-varying prior tunes the effect of the templates in different
regions of the reconstruction.

@@

\section{Application: Reconstruction for CT-guided radio-ablation study}
\label{sec:tmh}
Before diving into the details of the uniform and the
spatially-varying prior methods, we first show how both the
techniques can be applied to our advantage in a real-life medical
longitudinal study. Here, our data~\footnote{Source: Tata Memorial
  Centre~\cite{tmh}, Parel, Mumbai.  This is the national
  comprehensive centre for the prevention, treatment, education and
  research in cancer, and is recognized as one of the leading cancer
  centres in India.} consists of successive scans of the liver taken
during a radio-frequency ablation procedure. In such a
procedure~\cite{Dong2015}, the physician inserts a thin needle-like
probe into the organ. Repeated CT scans of the patient need to be
acquired in order to track the movement of the needle and to ensure
that it is reaching the appropriate target tumor. Once the needle hits
the tumor, a high-frequency electric current is passed through the tip
of the probe and this burns the malignant tumor (ablation). The scan
at this ablation stage must also reveal accurate details of the
changes within. In this context, we classify the goal of any of our
reconstruction techniques into two categories: 
\begin{enumerate}
\item To track the position of the needle in a relatively well-reconstructed background.%Initially, when our goal is to only track the location of the probe while simultaneously reducing sub-sampling artefacts, we advocate capturing measurements from a small number of views (`few-view' imaging) and using uniform prior based reconstruction.
\item To accurately observe the new changes amidst a relatively well reconstructed background after the needle touches the tumor.%Later, when the probe is proximal to or touches the tumor, our goal is to observe details of tumor ablation while simultaneously reducing sub-sampling artefacts. In this case, we propose to capture measurements from a slightly higher number of views (`moderate-view' imaging) to acquire more information, and use a spatially-varying prior-based reconstruction. The spatially-varying technique moderates the effect of the prior in the reconstruction of new changes in the object being scanned.
\end{enumerate}

%--------------------------Results on TMH data--------------------------
%------------------RFA 2 data
\begin{figure}[h!]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template1.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 1}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template2.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 2}
    \end{subfigure}
     \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template3.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 3}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template4.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 4}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template5.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 5}
    \end{subfigure}
              \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template6_color.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 6}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/template7.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 7}
    \end{subfigure}
       \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/tmh/RFA2/colorTestIm_ablation.png}
%\captionsetup{labelformat=empty}       
 \caption{slice 8}
    \end{subfigure}     
     \caption{Radio-frequency ablation dataset: One of the slices (512$\times$512) from each of the 8 scan volumes of a longitudinal study dataset of the liver. Note that in volumes (a) through (g), the needle (shown in red in (f)) approaches the target tumor. (h) the organ after the ablation: this slice is displayed on a separate intensity scale to enable proper viewing of the region marked in green that shows the after-effects of ablating the tumor.} 
\label{fig:RFA2_test_object-prior}
\end{figure}

%--------------------RFA 2 few views
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{white}{green}{\includegraphics[width=0.23\columnwidth]{../images/tmh/RFA2/very_few_views/colorTestIm.png}}}\hfill
\subcaptionbox{FBP}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/fbp.png}}\hfill
\subcaptionbox{l1-ls}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/cs.png}}\hfill
\subcaptionbox{Proposed\\ method: uniform prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/plain_pca.png}}
\caption[Representative results-1]{\textbf{Goal: Track new changes.}\small{ Reconstruction of slice 7 (`test') of Fig.~\ref{fig:RFA2_test_object-prior} from only 90 views, using (b) FBP and no prior resulting in streaks, SSIM $= 0.48$ (c) l1-ls resulting in blurred bone structures, SSIM $=0.35$ and (d) uniform prior (slices 1-6 of Fig.~\ref{fig:RFA2_test_object-prior} are used as object-prior) resulting in clear bone structures with less streaks, SSIM $=0.55$. The region enclosed in red rectangle is our RoI as it contains both the new position of the needle and some background. All SSIM values are computed for this RoI.}}
\label{fig:RFA2_very_few_views}
\end{figure}

@@

\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{white}{blue}{\includegraphics[width=0.23\columnwidth]{../images/tmh/RFA2/few_views/colorTestIm.png}}}\hfill
\subcaptionbox{FBP}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/fbp.png}}\hfill
\subcaptionbox{l1-ls}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/cs_dct.png}}\hfill
\subcaptionbox{Proposed\\ method: uniform\\ prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/plain_pca.png}}
\subcaptionbox{Proposed\\ method: spatially-varying prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/weighted_pca_all_methods_kk_0_01.png}}
\subcaptionbox{Weights map}{\includegraphics[width=0.275\columnwidth]{../images/tmh/RFA2/few_views/weights_colormap.png}}
\caption[Representative results-2]{\textbf{Goal: Observe details of new changes.} Reconstruction of slice 8 of Fig.~\ref{fig:RFA2_test_object-prior} from 120 views, using (b) FBP with SSIM $= 0.50$ (c) l1-ls with SSIM $=0.46$ and (d) uniform prior, with SSIM $=0.51$ (\emph{notice dominance of the prior: a prominent residual shadow of the needle which was present in the object-prior, but not present in the test image}), and (e) spatially-varying prior with SSIM $=0.56$ (\emph{notice that the dominance of the prior is significantly controlled}). The region enclosed in green rectangle is our Region of Interest (RoI) as it contains both the new position of the needle and some background. The SSIM is computed in this RoI. (f) shows the computed weights map (defined later in the paper) used for reconstruction. Darker intensities indicate lower weights to prior as these are the regions of new changes.}
\label{fig:RFA2_few_views}
\end{figure}

%Here, we demonstrate the combined use of l1-ls (CS) with the object
%prior, in two flavors: the vanilla (uniform) and spatially-varying
%prior-based reconstruction.
The choice of the number of
measurement views and the type of reconstruction -- uniform or
spatially-varying, is driven by the goal of the procedure. First, in
order to track the needle, a very small number of views is sufficient
because the needle has a very high attenuation coefficient when
compared to that of the organs. We use the uniform prior
reconstruction here to reduce the artefacts due to sub-sampling. The
uniform method is fast and sufficient to track the position of the
needle.  Once the needle reaches the site of the tumor, we propose
changing the imaging protocol to acquire measurements from a moderate
number of views. This will enable us to get more information about the
new changes. In addition, we then deploy the spatially-varying prior
method in order to locate the regions of new changes and penalize any
dominance of the prior in these regions.  Regardless of the imaging
protocol we use (`few' or `moderate'), the number of views is smaller
(atleast one-fifth) than the conventional number of views used in a
standard hospital setting.

The dataset from this longitudinal medical study consists of 8 scans
taken during the ablation procedure. We demonstrate our method for 2D
reconstruction by choosing a single slice from each of the 8 volumes
as our dataset. Note that all these 8 slices are located at the
\emph{same} index~\footnote{The notion of~\textit{same index} (slice
  number corresponding to the same depth) makes sense in the context,
  because in such problems, the different scans are aligned with each
  other.} within each of the respective
volumes. Fig.~\ref{fig:RFA2_test_object-prior} shows the chosen set of
2D slices (each of size 512$\times$512) from the different
volumes. Observe that the needle is seen in all of the first 7 slices
and the effect of ablation is seen in the 8th slice. \\

@@

\textbf{Observing details of the ablation:} Next, we choose slices 1-7 as our object-prior and reconstruct slice 8 from 120 views i.e., a somewhat higher number of views this time. Fig.~\ref{fig:RFA2_few_views} shows the reconstructions of slice 8 by different methods. We see that the spatially-varying prior reconstruction brings in the advantage of the prior without adversely affecting the reconstruction of new regions.

%---------------------------------Uniform section---------------------
\section{Uniform prior-based reconstruction}
\label{sec:method_uniform_prior}
Having presented the application, we first review the algorithm~\cite{my_dicta_paper} for a uniform  prior-based reconstruction. Principal Component Analysis (PCA) has been traditionally used to find the significant modes of (Gaussian corrupted) data. In this regard, it has been widely applied in the context of data compression. However, PCA can also be seen as a tool to provide an orthogonal basis to represent the space in which most of the test data could lie (except the new changes). This space is constructed from the available set of previously scanned objects which must cover a realistically representative range of structures. %, which was shown~\cite{my_dicta_paper} to be better when compared to dictionary-based priors. %A preliminary version of this particular method was earlier presented in \cite{my_dicta_paper}.

We first present the eigenspace-cum-CS prior-based reconstruction. To begin with, when an object is scanned multiple times, a set of high quality reconstructions (i.e., reconstructions from a dense set of projection views) may be chosen as object-prior for the reconstruction of future scan volumes, which in turn, may be scanned using far fewer measurements. The eigenspace $E_{\text{high}}$ of the $L$ previously scanned objects $Q_1,Q_2,...,Q_L$ is pre-computed. Here, it is assumed that the test volume, barring the new changes, can be expressed as a sparse linear combination of the principal components (eigenvectors of the covariance matrix) obtained from a group of structurally similar volumes. Hence, the object-prior is represented by means of PCA. For the eigenspace to encompass a range of possible structures in the test slice, the object-prior must represent a wide structural range. Moreover, if these volumes are not aligned, then they must be first registered before computing the prior. The prior is built by computing the covariance matrix from the template set $\{Q_i\}_{i=1}^L$. The space spanned by the eigenvectors $\{\boldsymbol{V_k}\}_{k=1}^{L-1}$ (eigenspace) of the covariance matrix is the object prior and is assumed to contain most of the test slice that is similar, but not necessarily identical to the object-prior. We use all of the $L-1$ orthogonal eigenvectors as a basis to represent the unknown test volume. Let $\boldsymbol{\mu}$ denote the mean of the previously scanned objects, and $\boldsymbol{\alpha}$ the vector of eigen-coefficients of the test scan, of which $\alpha_k$ is the $k^{\textrm{th}}$ element. Then, once the eigenspace is pre-computed, the test is reconstructed by minimizing the following cost function:
\vspace{4mm}
 \begin{equation}
   \setlength{\belowdisplayskip}{0pt} \setlength{\belowdisplayshortskip}{0pt}
\setlength{\abovedisplayskip}{-2pt} \setlength{\abovedisplayshortskip}{-2pt}
J_1(\boldsymbol{\theta,\alpha}) = \lVert\boldsymbol{\mathcal{R} x}- \boldsymbol{y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1+\lambda_2\lVert\boldsymbol{x} - (\boldsymbol{\mu} + \sum_{k}\boldsymbol{V_k}\alpha_k)\rVert_2^2.
\label{Eq:main_prior}
\end{equation}
Here, $\lambda_1$, $\lambda_2$ are tunable weights given to the sparsity and prior terms respectively. The unknowns $\boldsymbol{\theta}$ and $\boldsymbol{\alpha}$
are solved by alternately minimizing $J_{\boldsymbol{\alpha}}(\boldsymbol{\theta})$ using a fixed $\boldsymbol{\alpha}$, and $J_{\boldsymbol\theta}(\boldsymbol{\alpha})$ using the resultant $\boldsymbol{\theta}$, where 
\begin{align}
J_{\boldsymbol{\alpha}}(\boldsymbol{\theta}) &\triangleq \lVert\boldsymbol{\mathcal{R} x- y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1+\lambda_2\lVert\boldsymbol{x} - (\boldsymbol{\mu + V\alpha})\rVert_2^2, \\
J_{\boldsymbol\theta}(\boldsymbol{\alpha}) &\triangleq \lVert\boldsymbol{\Upsilon\theta} - (\boldsymbol{\mu + V\alpha})\rVert_2^2.
\end{align}
Note that $\boldsymbol{\theta}$ is solved for using the basis pursuit CS solver~\cite{l1ls}. Solving for $\boldsymbol{\alpha}$ leads to the closed form update:
\begin{equation}
\boldsymbol{\boldsymbol{\alpha}} = \boldsymbol{V}^T(\boldsymbol{\Upsilon \theta} -\boldsymbol{\mu}).
\end{equation}
 Optimal values of $\lambda_1, \lambda_2$ must be empirically chosen \textit{a priori}, based on the reconstructions of one of the template volumes (see also Sec.~\ref{sec:discussion}). The cost function described in Eq.~\ref{Eq:main_prior} is biconvex and the convergence of this optimization is guaranteed by the monotone convergence theorem~\cite{monotone}.
 %---------------------------------------------------------
 \vspace{2mm}
